---
title: "Exercise 12: Bootstrapping and Tree Support"
author: "Ruth Walters"
format: html
page-layout: full
---

```{r}
library(phangorn)
source("~/Documents/phylo/Exercise10/reRoot.R")
```

```{r}
primates.dna <- read.dna("~/Documents/phylo/Exercise5/data/primates.raw.aligned.fasta",
                         format = "fasta")

primates.dna
```

```{r}
primates.phyDat <- phyDat(primates.dna, 
                          type = "DNA", 
                          levels = NULL)

primates.phyDat
```

```{r}
mt <- modelTest(primates.phyDat, 
                model = "all")

mt
```

```{r}
best.model <- mt[which.min(mt$AIC),]

best.model
```

**Question 1**: Yes

```{r}
env <- attr(mt, "env")

best.fit <- eval(
  get(
    as.character(
      best.model[1]), 
    env), 
  env)

best.fit
```

```{r}
rooted.tree <- reroot.tree(best.fit$tree, 
                           "ringtailed_lemur")

plot(rooted.tree)
```

```{r}
bs <- bootstrap.pml(best.fit,
                    bs = 100, 
                    optNni = TRUE,
                    contro = pml.control(trace = 0))
```

```{r}
plotBS(rooted.tree, 
       bs, 
       p = 0.50, 
       type = "p")
```

**Question 3**: The best supported relationship (0.98) is at the node separating (humans + chimpanzees) and gorillas. The rest of the nodes have moderate but not "good" support (more than 0.60 but lower than 0.70).

**Question 4**: There is some evidence that humans and chimps are sister, but the relationship still does not meet the threshold for being considered "well-supported"

**Question 5**: Everyone has a different threshold for what they'll consider "support", and how that translates into an actionable conclusion. While the relationship between humans and chimps doesn't quite meet the 0.70 bar, it *is* close enough for me to say "probably". 
 