---
title: "Parsimony Exercise"
output:
  html_document: default
  pdf_document: default
date: "2026-02-03"
author: Ruth Walters
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(phangorn)
```

```{r}
primate <- read.delim("data/data.txt")
```

```{r}
primate
```


```{r}
primate_df <- data.frame(primate[,2:11], row.names = primate[,1])
primate.mat <- as.matrix(primate_df)
levels <- c("0", "1", "2", "3")
primate.phy <- phyDat(primate.mat, type = "USER", levels = levels)
primate.phy
```

```{r}
dm <- dist.hamming(primate.phy)
treeUPGMA <- upgma(dm)
```

```{r}
tree_parsimony <- optim.parsimony(treeUPGMA, primate.phy)
```

**Question 1**: It runs very quickly (< 1 second). Returns a p-score of 21 after 1 nni operation. 

```{r}
plot.phylo(tree_parsimony)
```

**Question 3**: The root taxon has three branches (it is not properly bifurcated).

```{r}
tree_parsimony <- root(tree_parsimony, 
                       outgroup = "ringtailed_lemur", 
                       resolve.root = TRUE)

plot.phylo(tree_parsimony)
```

**Question 5**: Rhesus monkey + chimpanzee + gorilla form a monophlyetic clade. All the organisms in this group have:

-  a protruding jaw, enlongated canines, a prominent orbital ridge/crest, a bony step-off at the back of the skull
- a similar number of cranial vein holes in their skull
- orbital diameter
- number of teeth and pointy molars

Among these traits, the ones that are not shared by other organisms (suggesting a synaptomorphy) include the protruding jaw.

