---
title: "Tutorial 1: R Syntax"
output: html_document
date: "2026-01-20"
---

## Installling packages

```{r}
library(ape)
```

## Functions

```{r}
random_sample <- runif(n = 10, min = 1, max = 2)
```

The line above is a call to the function `runif`. We know `runif` is a function because it is followed by parentheses. The statements inside it `n = 10`, `min = 1`, `max = 2` are arguments for the function. We know these are arguments because they are inside the parentheses and separated by commas. Finally, the arrow `<-` is used to assign the output (or value) of the function to the object `random_sample`.

The object `random_sample` will save the output from the function `runif` such that we can work with it later.


To check what `runif` just did we can read its help page:

```{r}
help( runif )
```

The help frame in the right now shows the help page for the `runif` function as well as other functions in the same family. The help page describes what the function computes and also the function arguments.

**Question 1**: Based on the help page, what `runif` does? 

**Question 2**: What are the meaning of the arguments we changed?

We can check what is the value of an object by printing it to the console screen:

```{r}
random_sample
```

Or using a syntax more familiar to other programming languages:

```{r}
print( random_sample )
```

Both formats will have the same outcome. Calling the name of a object will print its value to the screen.

**Question 3**: Is there any difference in the numbers printed to the screen when called 'random_sample' or 'print(random_sample)' above? Why?

Now, what do you think will happen if we define 'random_sample' again using the same function call to `runif`?

```{r}
random_sample <- runif(n = 10, min = 1, max = 2)
print(random_sample)
```

**Question 4**: Are the values stored on `random_sample` the same as before? Why?

## The anatomy of a command line

We will use lines of code to communicate with the R console. Although this seems annoying at first. A script with lines of code is a very powerful and reliable tool. Each line of code follow some syntax rules. These rules are very general.


Let's evaluate a function call and explain each piece: 

```{r}
sequence <- seq(from = 1, to = 5)
```


These are the elements of this line of code, in order:

- object_name 
- assign_symbol 
- function_call 
- open_parenthesis 
- argument_name 
- equal_sign 
- argument_value 
- comma 
- argument_name 
- equal_sign 
- argument_value 
- close_parenthesis

Many errors in R happen because:

1) You forgot to open or close the parentheses.
2) You used a wrong name for an argument.
3) You forgot the comma separating arguments.

It is important to pay attention to the basic syntax of R. 

