---
title: "Assessment"
format: pdf
editor: visual
---

```{r}
#| message: false
#| warning: false
library(phangorn)
library(msa)
source("assessRoot.R")
```

```{r}
viral_dna <- read.dna("RidgeFever_23GENOMES.fasta",
                      format = "fasta")

viral_dna
```

```{r}
viral_phy <- phyDat(viral_dna, 
                    type = "DNA", 
                    levels = NULL)

viral_phy
```

```{r}
mt <- modelTest(viral_phy, 
                model = "all")

best.model <- mt[which.min(mt$AIC),]

best.model
```

```{r}
env <- attr(mt, "env")

best.fit <- eval(
  get(
    as.character(
      best.model[1]), 
    env), 
  env)

best.fit
```

```{r}
#| fig-height: 12
#| fig-width: 18
rooted.tree <- reroot.tree(best.fit$tree)

plot(rooted.tree)
```

```{r}
bs <- bootstrap.pml(best.fit,
                    bs = 100, 
                    optNni = TRUE,
                    contro = pml.control(trace = 0))
```

**Question 1**: See plot:

```{r}
#| fig-height: 18
#| fig-width: 12
plotBS(rooted.tree, 
       bs, 
       p = 0.50, 
       type = "p",
       align.tip.label = TRUE, 
       bs.adj = 1.5)
```

**Question 2**: The Human RFV genomes form a clade that is sister to the Bat Ridge-Like virus, which shares a recent common ancestor with the Raccoon Ridge Virus and the Fox Ridge Virus. This close relationship with animal viruses suggests that the *wildlife spillover hypothesis* is the most correct.

**Question 3**: The support for this tree is generally very high, and the bootstrap value for the Human RFV branch meets the standard for evidence (0.73 \> 0.70). This suggests that the relationships are supported by genetic data.
